Empirical normative theories

### 2020-06-24 simple.tex
Moved draft section

Given any $(\alltopic, ¬, \allargs, I)$, a \emph{normative claim} about $(\alltopic, ¬, \allargs, I)$ (or about $\allargs$, when the other defined objects are left implicit) is any first-order logic proposition involving the atoms $[(\hist, \phi) \in {\ileadstoe[i]}], \hist \in \allhist, \phi \in \Phi$ which, together with \cref{ax:nocontr,ax:norm}, permits to deduce that $\phi \in \iPhi$ for some $i \in I$ and $\phi \in \Phi$. The claim is said to be \emph{about} the given set $\allargs$. An observation protocol, together with observables $(\ileadstoe[i])_{i \in I}$, \emph{falsifies} a normative claim iff the falsity of the claim is deducible from \cref{ax:nocontr,ax:norm} and the observations. A normative claim is \emph{falsifiable} iff there exists a finite observation protocol such that for any observables $(\ileadstoe[i])_{i \in I} \in (\allhist × \Phi)^I$, if the claim is false given these observables, then it is possible to deduce it from \cref{ax:nocontr,ax:norm} and the observations given the observables.

\begin{proposition}[Trivial claims are not falsifiable]
	Given any $(\alltopic, ¬, \allargs ≠ \emptyset, I)$, no minimal normative claim is falsifiable.
\end{proposition}
\begin{proof}
	Consider any claim about some $i$ and some $\phi$ and any finite observation protocol. We have to show that for some observables, the claim is false, but it is impossible to observe it (meaning, to deduce it from \cref{ax:nocontr,ax:norm} and observations).

\begin{theorem}
	If an empirical theory is convincing, and its normative theory satisfies \cref{ax:nocontr,ax:norm}, then it is valid.
\end{theorem}
\begin{proof}
	Consider $i \in I$ and $\phi \in {\gleadsto}(\allargs)$. 
	By hypothesis, $\exists k$ such that some $\ar_k \in \args_k$ defends $\phi$ under protection of $\emptyset$. Therefore, $\ar_k \ileadstosts \phi$. Using \cref{ax:norm}, this yields that $\phi \in \iPhi$, and validity follows from \cref{ax:nocontr}.
\end{proof}

\begin{theorem}
	The convincingness claim of any empirical theory is falsifiable: given an ET, there exists a finite obs prot with memory such that for any observables, if the claim is false given these observables, then it is possible to deduce it from the axioms and observations.
\end{theorem}
\begin{proof}

\end{proof}

\section{Additive empirical theories}
TODO: 3) define single-answer models. 4) phrase conditions below for history and single-answer models. 5) show that these conditions can be verified or falsified. 6) show that if the claim of the model does not include these conditions, then the model is not falsifiable, as suggested here below for some form of claims.

\subsection{Definition}
An empirical theory for a normative theory is a tuple $(q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, +)$,
with ${\gleadsto[q(i)]} \subseteq \allargs × \Phi$, ${\gbeats} \subseteq \allargs × \allargs$,
${\gbeatsinv}(\allargs)$ finite,
and $+$ a commutative associative internal binary operation on $\allargs$. Given a finite $\args \subseteq \allargs$, the notation $\sum \args = \sum_{\ar \in \args} \ar$ thus designates a well-defined argument. 
\commentOC{Check about finiteness.}
\commentOC{${\gbeatsinv}$ need not be finite, only those even ones: the transitive closure of ${\gleadstoinv}(\Phi)$ through ${\gbeatsinv}^2$, that is, $({\gbeatsinv})^2[{\gleadstoinv}(\Phi)] = \cup_{k \in 2\N}({\gbeatsinv}^{k}({\gleadstoinv}(\Phi)))$.}
\commentOC{I should only define $\ar \in \allargs$ plus $\ard$.}

@@

>>>>

\section{Additive single-answer empirical theories}
\subsection{Definition}
An empirical theory for a normative theory is a tuple $(q, (\gleadsto[q(i)])_{q(i) \in q(I)}, \gbeats, +)$,
with ${\gleadsto[q(i)]} \subseteq \allargs × \Phi$, ${\gbeats} \subseteq \allargs × \allargs$,
${\gbeatsinv}(\allargs)$ finite,
and $+$ a commutative associative internal binary operation on $\allargs$. Given a finite $\args \subseteq \allargs$, the notation $\sum \args = \sum_{\ar \in \args} \ar$ thus designates a well-defined argument. 
\commentOC{Check about finiteness.}
\commentOC{${\gbeatsinv}$ need not be finite, only those even ones: the transitive closure of ${\gleadstoinv}(\Phi)$ through ${\gbeatsinv}^2$, that is, $({\gbeatsinv})^2[{\gleadstoinv}(\Phi)] = \cup_{k \in 2\N}({\gbeatsinv}^{k}({\gleadstoinv}(\Phi)))$.}
\commentOC{I should only define $\ar \in \allargs$ plus $\ard$.}

@@

Show that if the claim of the model is of a different form, then the model is not falsifiable, as suggested here below for some form of claims.

Extend the definition of additive single-answer theories to non-additive single-answer theories.

Given any $(\alltopic, ¬, \allargs, I)$, a \emph{normative claim} about $(\alltopic, ¬, \allargs, I)$ (or about $\allargs$, when the other defined objects are left implicit) is any first-order logic proposition involving the atoms $[(\hist, \phi) \in {\ileadstoe[i]}], \hist \in \allhist, \phi \in \Phi$ which, together with the axioms, permits to deduce that $\phi \in \iPhi$ for some $i \in I$ and $\phi \in \Phi$. The claim is said to be \emph{about} the given set $\allargs$. An observation protocol, together with observables $(\ileadstoe[i])_{i \in I}$, \emph{falsifies} a normative claim iff the falsity of the claim is deducible from the axioms and the observations. A normative claim is \emph{falsifiable} iff there exists a finite observation protocol such that for any observables $(\ileadstoe[i])_{i \in I} \in (\allhist × \Phi)^I$, if the claim is false given these observables, then it is possible to deduce it from the axioms and the observations given the observables.

\begin{proposition}[Trivial claims are not falsifiable]
	Given any $(\alltopic, ¬, \allargs ≠ \emptyset, I)$, no minimal normative claim is falsifiable.
\end{proposition}
\begin{proof}
	Consider any claim about some $i$ and some $\phi$ and any finite observation protocol. We have to show that for some observables, the claim is false, but it is impossible to observe it (meaning, to deduce it from the axioms and observations).

\begin{theorem}
	If an empirical theory is convincing, and its normative theory satisfies the axioms, then it is valid.
\end{theorem}
\begin{proof}
	Consider $i \in I$ and $\phi \in {\gleadsto}(\allargs)$. 
	By hypothesis, $\exists k$ such that some $\ar_k \in \args_k$ defends $\phi$ under protection of $\emptyset$. Therefore, $\ar_k \ileadstosts \phi$. Using \cref{ax:norm}, validity follows.
\end{proof}

\begin{theorem}
	The convincingness claim of any empirical theory is falsifiable: given an ET, there exists a finite obs prot with memory such that for any observables, if the claim is false given these observables, then it is possible to deduce it from the axioms and observations.
\end{theorem}
